Discussion Draft

### 2020-04-15 doc/Discussion.rmd
fixed wierd underline big, moved a comment out section

---
title: "Discussion Draft"
date: ''
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

@@

```{r Model.testing, include=FALSE}

#imortant of testable hypotheses

#model not included in this comparison

@@

0. Regardless of if these hypotheses turn out to be true, our results extend previous reports (dumont) of rapid male-specific evolution in multiple subspecies and strains of house mouse.

1. whether these instances of evolution are due to seperat independent events or segregating genetic variation within the population requires further study.

2. These results may hold importance to animal breeding programs -- may suggest that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most efficient way to increase the sex-averaged recombination rates. This could be important application for animal breeding programs (Battagin et al 2016).

@@

# consequence of greater female variance -- (between cell)
#

Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013]. <More details on the empirarcal pattern> (cite morelli cohen)

@@

<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big)review>

0. Reguardless of whether these models and hypothese are true ... if the pattern of females having more across cell variation in gwRR

@@

#(sardel kirkptarick -

# diviation from uniform is due to male rec landscape
# patterns of Fst (Hanel et al, but see Kartej?)
# source could be DSB formation along chromosome - and /or biases in DSB repair 
#CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick))  )
#

@@

<The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.>

If these hypotheses concerning evolution of these patterns **are not true** -- there's still predictions that can be made given the consequences of these patterns. The consequences of this difference in the typical landscape as reviewed in [@sardell_sex_2020]. For the sex averaged landscape along chromosomes, any diviation from uniform placement is generally due to male specific patterns.

-- There's also the prediction that females have active CO supression near telomeres 
(variation) -- geiven the evidence in (female mice) that show a higher elevation of DSBs at the telomere (halldorsson?).

@@

<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>

@@

A consequence of sexual dimorphism chromatin organization is that interference strength quantified in base pairs will be sexually dimorphic. There are two aspects that define this relationship. First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K)).  Second, the signal of crossover interference is mediated through the chromosome axis or SC (kleck Dernberg). Thus the chromatin organization with larger loops (and shorter axis) will have stronger crossover interference when quantified in base pairs, since more base pairs are sequestered in the chromatin loops per unit of axis or SC. Empirical data confirms that interference strength measured in micrometer along the axis is conserved between sexes while interference strength quantified from genetic map is sexually dimorphic, comes from mice [@petkov2007], humans (Gruhn, Lynn). However, due to lack of relevant cytology data in both sexes it is not clear if the many examples of sexual dimorphism in linkage maps is due to chromatin organization.

@@

<importance of evolutionary distinct trajectories, measure both sexes -- do evolutionary comparisons within sex.

@@

integrate physical models -- centromere and kinetochore
>

>>>>

---
title: "Discussion Draft"
date: ''
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '3'
bibliography: refs_Main_Report.bib
---

@@

@@

Regardless of if these hypotheses turn out to be true, our results extend previous reports (dumont) of rapid male-specific evolution in multiple subspecies and strains of house mouse.

whether these instances of evolution are due to seperat independent events or segregating genetic variation within the population requires further study.

These results may hold importance to animal breeding programs -- may suggest that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most efficient way to increase the sex-averaged recombination rates. This could be important application for animal breeding programs (Battagin et al 2016).

@@

# consequence of greater female variance -- (between cell)
```

Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013].

@@

<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromosomes, Mogessie et al 2018 -- meiotic spindle (big) review>

0. Reguardless of whether these models and hypothese are true if the pattern of females having more across cell variation in gwRR

@@

#sardel kirkptarick -

# diviation from uniform is due to male rec landscape
# patterns of Fst (Hanel et al, but see Kartej?)
# source could be DSB formation along chromosome - and /or biases in DSB repair 
#CO suppresion, elevated DSB (most repaired as NCO) -- elevated DsB (Brick)

@@

<The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I>

If these hypotheses concerning evolution of these patterns **are not true**  there's still predictions that can be made given the consequences of these patterns. The consequences of this difference in the typical landscape as reviewed in [@sardell_sex_2020]. For the sex averaged landscape along chromosomes, any diviation from uniform placement is generally due to male specific patterns.

There's also the prediction that females have active CO supression near telomeres (variation) -- geiven the evidence in (female mice) that show a higher elevation of DSBs at the telomere (halldorsson?).

@@

<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]>

@@

A consequence of sexual dimorphism chromatin organization is that interference strength quantified in base pairs will be sexually dimorphic. There are two aspects that define this relationship. First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K).  Second, the signal of crossover interference is mediated through the chromosome axis or SC (kleck Dernberg). Thus the chromatin organization with larger loops (and shorter axis) will have stronger crossover interference when quantified in base pairs, since more base pairs are sequestered in the chromatin loops per unit of axis or SC. Empirical data confirms that interference strength measured in micrometer along the axis is conserved between sexes while interference strength quantified from genetic map is sexually dimorphic, comes from mice [@petkov2007], humans (Gruhn, Lynn). However, due to lack of relevant cytology data in both sexes it is not clear if the many examples of sexual dimorphism in linkage maps is due to chromatin organization.

@@

importance of evolutionary distinct trajectories, measure both sexes -- do evolutionary comparisons within sex.

@@

integrate physical models -- centromere and kinetochore
